---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


# aiR

aiR is used to assess PM2.5 exposures in the Cincinnati, Ohio area. The package creates predictions based on a spatiotemporal hybrid satellite/land use random forest model. PM2.5 exposure predictions are available at 1 x 1 km grid resolution covering the "seven county" area (OH: Hamilton, Clermont, Warren, Butler; KY: Boone, Kenton, Campbell) on a daily basis from 2000 - 2015.

## Installing

aiR is a private package hosted on GitHub.

Install with:

```{r eval=FALSE}
remotes::install_github('cole-brokamp/aiR')
```


## Example

This example covers how to extract PM2.5 exposure estimates given latitude/longitude coordinates and dates.

Note that `pm_grid` and `pm_data` are both R objects that will be available upon loading of the package. However, `pm_data` has to be split into two smaller files to be under GitHub's 100 MB filesize limit.  This workaround requires binding the two datasets into one upon package loading.

```{r}
library(aiR)
pm_data <- bind_rows(pm_data_early, pm_data_late)
```


Create a demonstration dataset of random coordinates and dates:

```{r}
library(sf)
library(tidyverse)

d <- tibble::tribble(
  ~id,         ~lon,        ~lat,
    809089L, -84.69127387, 39.24710734,
    813233L, -84.47798287, 39.12005904,
    814881L, -84.47123583,  39.2631309,
    799697L, -84.41741798, 39.18541228,
    799698L, -84.41395064, 39.18322447
  )

set.seed(12)

d <- d %>%
    mutate(date = seq.Date(as.Date('2015-01-01'), as.Date('2015-12-31'), by=1) %>%
             base::sample(size=nrow(d)))
```

Convert this to a simple features object and transform to the Ohio South projection. Reprojection is necessary because `pm_grid` is in the Ohio South projection.

```{r example}
d <- d %>%
  st_as_sf(coords = c('lon', 'lat'), crs=4326) %>% 
    st_transform(3735)
```


To estimate the exposures, we will first overlay the locations with the PM2.5 exposure grid to generate the `pm_grid_id` for each location.

```{r}
( d <- st_join(d, pm_grid) )
```


Then merge the "lookup grid" (`pm_grid`) into the dataset by using `pm_grid_id` and `date`. Note that for the merge to work, the date column must be named `date` and be of class `Date` and the `pm_grid_id` column must exist.

```{r}
( d <- left_join(d, pm_data, by=c('pm_grid_id', 'date')) )
```


## Example adding weather data




